Adaptive Disk Storage for Interaction Graphs

### 2014-06-23 docs/graph-adapt/main.tex
authos
Added a more generalized overlapping attributes algorithm

@@

\textbf{Problem:} \emph{Find a true partitioning of attributes that minimizes
  the query I/O and bounds the storage cost by some upper limit.}

@@

\textbf{Problem:} \emph{Find an overlapping partitioning of attributes that minimizes
  the query I/O and bounds the storage cost by some upper limit.}

Below, we sketch two proposed algorithms. We are still working on these.

\paragraph{First proposal}

\begin{itemize}

\item First, let's define a measure for the similarity between two
partitions. I think this can be the sum of the size of attributes that
they do not have in common. So, if we have partitions: abc, abd, and
abe, then abc-abd will be more similar than abc-abe, if size(d) $<$
size(e).

\item The first partitioning is based on what queries we have seen.
Every query gets its own sub-block. This is the "ideal" situation,
because the I/O cost would be minimized for every query that we would
have seen.

\item We calculate the storage cost for the partitioning.

\item If the storage cost is above some threshold, then we do a pairwise
comparison of all sub-blocks to determine which two are the most
similar, using (0).

\item We combine the two most-similar sub-blocks, and repeat the process
starting at (2) until we are satisfied with the storage cost.

\item The result is our partitioned block.

\end{itemize}

\begin{algorithm}[h]
\scriptsize
\caption{Algorithm for partitioning blocks into sub-blocks with overlapping attributes.}
\label{alg:overlappingP}
\KwData{$B$: block, $Q$: set of queries}
$\mathcal{P}(B) \leftarrow \emptyset;$  \tcp*{Sub-blocks;}

@@

\While(\tcp*[f]{While storage exceeds threshold}){$H(\mathcal{P}, B) \geq \text{max}$}{
    $min = \infty$ \; 
    $x = \emptyset$ \;
    $y = \emptyset$ \;
    \For{$b_1 \in \mathcal{P}(B)$}{ 
        \For{$b_2 \in \mathcal{P}(B)$}{ 
            $\delta = (b_1 \cup b_2) - (b_1 \cap b_2)$ \; 
           \If{$min < \delta $}{
                 $\text{x} \leftarrow  b_1$\;
                 $\text{y} \leftarrow b_2$\;
                 $min \leftarrow \delta$
            } 
        }
    }
$ \mathcal{P}(B)  \leftarrow \mathcal{P}(B) - x - y$ \;
 $\mathcal{P}(B)  \leftarrow \mathcal{P}(B)  + x \cup y$ \;
}
\Return $ \mathcal{P}(B)$  \tcp*{Final set of sub-blocks}
\end{algorithm}

\paragraph{Second proposal}

\begin{itemize}

\item The first partitioning is based on what queries we have seen.
Every query gets its own sub-block. This is the "ideal" situation,
because the I/O cost would be minimized for every query that we would
have seen.

\item We keep a mapping of queries to blocks and frequencies.

\item If the storage cost is above some threshold, then we get the queries with
  the lowest frequencies, and combine their blocks. We also add the frequencies,
  so we are combining the blocks that are least likely to be accessed.

\item We repeat until we are below the storage threshold.

\item The result is our partitioned block.

\end{itemize}

\begin{algorithm}[h]
\scriptsize
\caption{Alternative algorithm for partitioning blocks into sub-blocks with overlapping attributes.}
\label{alg:overlappingP2}
\SetKwFunction{Enqueue}{Enqueue}
\SetKwFunction{Dequeue}{Dequeue}

\KwData{$B$: block, $Q$: set of queries}
$\mathcal{P}(B) \leftarrow \emptyset;$  \tcp*{Sub-blocks;}
$pq \leftarrow \emptyset;$  \tcp*{priority queue;}

\For(\tcp*[f]{Every query gets its own sub-block}){$q \in Q$}{ 
 $\mathcal{P}(B) = \mathcal{P}(B) \cup q.A$ \;
 \Enqueue{$pq,\mathcal{P}(B), f(\mathcal{P}(B))$}\tcp*{put blocks in the queue;}
}
\While(\tcp*[f]{While storage exceeds threshold}){$H(\mathcal{P}, B) \geq
  \text{max}$}{  
  $x \leftarrow $ \Dequeue{$pq$}\;
$y \leftarrow $ \Dequeue{$pq$}\tcp*{remove the two minimum;}
  \Enqueue{$pq,x \cup y, f(x) + f(y)$}\tcp*{combine and insert;}    
    }

\For(\tcp*[f]{The elements in the queue are our partition}){$p \in pq$}{ 
  $ \mathcal{P}(B)  \leftarrow \mathcal{P}(B) \cup p$ \;
}
\Return $ \mathcal{P}(B)$  \tcp*{Final set of sub-blocks}
\end{algorithm}

>>>>

\DeclareMathOperator{\dist}{dist}

@@

\paragraph*{Problem.}\emph{Find a true partitioning of attributes that minimizes
  the query I/O and bounds the storage cost by some upper limit.}

\paragraph*{Approach.}

@@

\paragraph*{Problem.}\emph{Find an overlapping partitioning of attributes that minimizes
  the query I/O and bounds the storage cost by some upper limit.}

\paragraph*{Approach.} We start the algorithm with a 
 partitioning based on what queries we have seen.
Every query gets its own sub-block. This is the``ideal'' partitioning,
because the I/O cost would be minimized for every query that we would
have seen. As input to the algorithm, we provide a function $\dist$, which
is used to compute the distance between two blocks.

$$
\dist: block \rightarrow block \rightarrow int
$$

Note that the $\dist$ function can be chosen arbitrarily.  Examples of useful
$\dist$ functions are the frequencies with which the queries are accessed, or the Jaccard index, which measures the set-similarity of
the attributes.

Using this function, the algorithm iteratively combines the two partitions that
are closest together. Algorithm~\ref{alg:closestBF} shows the pseudocode for a brute-force approach
  to finding the closest paris that runs in $O(n^2)$ time. However, there is
  also a divide-and-conquer algorithm~\cite{cormen01} for solving closest pair
  of points problem in $O(n \log n)$.
After each combination of partitions, the algorithm calculate the storage cost
for the partitioning. The algorithm stops when the  storage cost is below some
specified threshold.  The result is the block partitioning.

\begin{algorithm}[h]
\SetKwFunction{closestPair}{closestPair}
\scriptsize
\caption{Algorithm for partitioning blocks into sub-blocks with overlapping attributes.}
\label{alg:overlappingP}
\KwData{$B$: block, $Q$: set of queries}
$\mathcal{P}(B) \leftarrow \emptyset;$  \tcp*{Sub-blocks;}

@@

\While(\tcp*[f]{While storage exceeds threshold}){$H(\mathcal{P}, B) \geq
  \text{max}$}{
$\text{x}, \text{y}  \leftarrow  \closestPair{}  $\;
 $ \mathcal{P}(B)  \leftarrow \mathcal{P}(B) - x - y$ \;
  $\mathcal{P}(B)  \leftarrow \mathcal{P}(B)  + x \cup y$ \;
}
\Return $ \mathcal{P}(B)$  \tcp*{Final set of sub-blocks}
\end{algorithm}

 \begin{algorithm}[h]
\SetKwFunction{length}{length}
\SetKwFunction{dist}{dist}
 \scriptsize
 \caption{Brute-force algorithm for closest points in $O(n^2)$.}
 \label{alg:closestBF}

$minDist  \gets \infty $ \;
\For{$i \gets 1$ \textbf{to} $\length(P)-1$} {
\For{$j \gets i+ 1$ \textbf{to} $\length(P)$} {

  $p \gets P[i]$ \;
  $q \gets P[j]$ \;
  \If{$\dist(p,q) < minDist$} {
    $minDist \gets \dist(p,q)$\;
    $closestPair \gets (p,q)$\;
  }
}
}
\Return $closestPair$

% \KwData{$B$: block, $Q$: set of queries}

% $\mathcal{P}(B) \leftarrow \emptyset;$  \tcp*{Sub-blocks;}

% \For(\tcp*[f]{Every query gets its own sub-block}){$q \in Q$}{ 
%  $\mathcal{P}(B) = \mathcal{P}(B) \cup q.A$ 
% }

% \While(\tcp*[f]{While storage exceeds threshold}){$H(\mathcal{P}, B) \geq
%   \text{max}$}{
% $\text{x}, \text{y}  \leftarrow  \closestPair{}  $\;
%  $ \mathcal{P}(B)  \leftarrow \mathcal{P}(B) - x - y$ \;
%   $\mathcal{P}(B)  \leftarrow \mathcal{P}(B)  + x \cup y$ \;

% \Return $ \mathcal{P}(B)$  \tcp*{Final set of sub-blocks}
 \end{algorithm}
